---
title: "Stacker News"
subtitle: "An analysis of a payment-based online forum"
author: "Alberto & Filippo"
date: "2023-10-27"
output: 
  html_document: 
    toc: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = "UTF-8", output_dir = "../outputs/")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(remotes)
#remotes::install_github("edgararuiz/connections")
library(connections)
library(RSQLite)
library(data.table)
library(igraph)
library(ggplot2)
library(stringr)
```

## Data importing, cleaning and manipulation

```{r Query to database}
conn <- connection_open(SQLite(), "../data/stacker_news.sqlite")

users <- setDT(dbGetQuery(conn, "SELECT * FROM user"))
comments <- setDT(dbGetQuery(conn, "SELECT * FROM comments"))
posts <- setDT(dbGetQuery(conn, "SELECT * FROM post"))

connection_close(conn)
```

Please note that while scraping data were all saved in 'string` form to better
manage datatypes in the database

### Users

```{r}
str(users)
```

Users table contains five variables.

- `User` is a string, the username;
- `TotalStacked` is the total amount of sat ever stacked by the user -> need it as integer;
- `FirstItem` is the number of the first post that the user created, if any;
- `HatStreak` is the number of cowboy hat streaks that the user achieved, if any -> need it as integer;
- `NumItems` is the total number of items created by the user in the forum, if any -> need it as integer.

The following chunck changes the needed variables into the needed datatype

```{r}
change_columns <- c("TotalStacked", "HatStreak", "NumItems")  
users[ , (change_columns) := lapply(.SD, as.numeric), .SDcols = change_columns]
```

Result of the transformation

```{r}
sapply(users, class)
```

### Posts

```{r}
str(posts)
```

Post table contains 13 variables.

- `Title` is the post title;
- `Category` is the post category;
- `ItemCode` is the unique identifier for the post;
- `Sats` is the number of sat stacked by the post -> need it as integer;
- `Boost` is the number of boosts assigned to the post. Boosts are needed to speed up the post diffusion in the form (kind of advertising) -> need it as integer;
- `Comments` is the number of comments -> need it as integer;
- `Author` is the username of the post author;
- `Tag` is the tag assigned to the post by the author;
- `Timestamp` is the timestamp of the creation of the post -> need it in datetime;
- `MainLink` is the main link displayed by the post, a post has a main link if it has the *link* category, if not then the value is NA;
- `BodyLinks` is the field containing a string in which all the links in the body post and comments are contained. If the post contains no link, then the string is empty;
- `SatsReceivedComments` is the total number of sat received by the comments to the post -> need it as integer;
- `CommentsItemCode` is the list of item codes of the comments.

Multiple variables of this table must be transformed into the proper data type, however some of those cannot be transformed straightforwardly because of multiple formatting errors. A more profund data cleaning is needed before the datatype transformation.

#### Sats variable

This variable has to be transformed into numeric type, however beforehand we must cut away the postfix 'sats' or 'sat' from the stings.

```{r}
# Substitute rounding of thousands 'k' and millions 'm' into integers
posts[ , Sats := (str_replace_all(Sats, "k sats$", "000") %>%
                    str_replace_all(., "m sats$", "000000") %>%
                    str_replace_all(., " sats$", "") %>%
                    str_replace_all(., " sat$", "")
                  )
       ]

posts[, Sats := ifelse(str_detect(Sats, "\\."), str_sub(str_replace_all(Sats, "\\.", ""), end = -2), Sats)]

# Turn variable in numeric
posts[ , Sats := as.numeric(Sats)] # The rows with sat amount equal to 'None' will inherit an NA value
```

Please note that we are left some NAs. These items with Sats=NA are items that don't have the Sats field in the post template because they are created by the forum administrator and they cannot be tipped.
Also 37 of them are created by the user 'ad', who is a bot defined by the forum creator to broadcast ads as posts. Those items cannot be tipped too.

```{r}
# Test if we managed None values correctly
nrow(posts[is.na(Sats)]) == nrow(posts[is.na(Sats) & Author=='saloon']) + nrow(posts[is.na(Sats) & Author=='ad'])
```

Now the `Sats` column contains only numeric values (the amount of sat stacked by the post) and the NAs are all related to posts that actually didn't receive Sats because of structural reasons.

The following graph shows the distribution of sats within the posts. The distribution is clearly asymmetric and has the majority of the posts with an amount of sats stacked very close to zero.

```{r}
posts %>%
  ggplot()+
  geom_density(aes(x = Sats))
```

By filtering the amount of sats to a reasonable value is possible to observe the distribution, with the median value of sats stacked by a post equal to `r na.omit(posts, "Sats")[, median(Sats)]`

```{r}
posts[Sats<1000] %>%
  ggplot()+
  geom_histogram(aes(x = Sats))
```


#### Boost variable















