---
title: "Untitled"
author: "Alberto Bersan 879597"
date: "2023-11-02"
output:
  html_document:
    toc: true
    toc_depth: 3
---
# Library
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(remotes)
#remotes::install_github("edgararuiz/connections")
library(connections)
library(RSQLite)
library(data.table)
library(igraph)
library(ggplot2)
library(stringr)
library(lubridate)
library(tidyverse)
```

# Data Import
```{r Query to database}
conn <- connection_open(SQLite(), "stacker_news.sqlite")
users <- setDT(dbGetQuery(conn, "SELECT * FROM user"))
comments <- setDT(dbGetQuery(conn, "SELECT * FROM comments"))
posts <- setDT(dbGetQuery(conn, "SELECT * FROM post"))

connection_close(conn)
```

# PART 1. Data Cleaning
```{r}
change_columns <- c("TotalStacked", "HatStreak", "NumItems")  
users[ , (change_columns) := lapply(.SD, as.numeric), .SDcols = change_columns]
sapply(users, class)
```

### users label
```{r}
contrib <- c("k00b",
                  "kr",
                  "ekzyis",
                  "WeAreAllSatoshi",
                  "rleed",
                  "bitcoinplebdev",
                  "benthecarman"
                  )

users[, Role := ifelse(User %in% contrib, "contributor", "other")]
```

### Sats and Boost k 
```{r}
# Substitute rounding of thousands 'k' and millions 'm' into integers (sats)
posts[ , Sats := (str_replace_all(Sats, "k sats$", "000") %>%
                    str_replace_all(., "m sats$", "000000") %>%
                    str_replace_all(., " sats$", "") %>%
                    str_replace_all(., " sat$", "")
                  )
       ]

posts[, Sats := ifelse(str_detect(Sats, "\\."), str_sub(str_replace_all(Sats, "\\.", ""), end = -2), Sats)]

# Turn variable in numeric
posts[ , Sats := as.numeric(Sats)] 

# Substitute rounding of thousands 'k' (boost)
posts[ , Boost := (str_replace_all(Boost, "k boost$", "000") %>%
                    str_replace_all(., " boost$", "")
                  )
       ]

posts[, Boost := ifelse(str_detect(Boost, "\\."), str_sub(str_replace_all(Boost, "\\.", ""), end = -2), Boost)]

posts[, Boost := as.numeric(Boost)]
```

Test if we managed None values correctly
```{r}

nrow(posts[is.na(Sats)]) == nrow(posts[is.na(Sats) & Author=='saloon']) + nrow(posts[is.na(Sats) & Author=='ad'])
```

### Timestamp variable
```{r}
posts[, Timestamp := ymd_hms(Timestamp)]
```

### Bodylink
Since BodyLinks values are in the string form, we need firstly to deal with BodyLinks datatype, turning it into a vector. Then we can execute the needed procedure.
```{r}
posts[, BodyLinks := (str_replace_all(BodyLinks, '^\\[', "") %>%
                        str_replace_all(., '\\]$', "")
                      )
      ]

posts[, BodyLinks := list(strsplit(BodyLinks, split = ", "))]
```

The following code then extract the first link from the BodyLinks column and put it in the MainLink column, for posts that are links.
```{r}
posts[, MainLink := ifelse(MainLink=='None', sapply(BodyLinks, `[`, 1), MainLink)]
```

### SatsReceivedComments variable
```{r}
posts[ , SatsReceivedComments := (str_replace_all(SatsReceivedComments, "k sats$", "000") %>%
                                    str_replace_all(., "m sats$", "000000") %>%
                                    str_replace_all(., " sats$", "") %>%
                                    str_replace_all(., " sat$", "")
                                  )
       ]

posts[, SatsReceivedComments := ifelse(str_detect(SatsReceivedComments, "\\."), str_sub(str_replace_all(SatsReceivedComments, "\\.", ""), end = -2), SatsReceivedComments)]

posts[, SatsReceivedComments := as.numeric(SatsReceivedComments)]
```

### Sats variable
```{r}
comments[ , Sats := (str_replace_all(Sats, "k sats$", "000") %>%
                    str_replace_all(., "m sats$", "000000") %>%
                    str_replace_all(., " sats$", "") %>%
                    str_replace_all(., " sat$", "")
                  )
       ]

comments[, Sats := ifelse(str_detect(Sats, "\\."), str_sub(str_replace_all(Sats, "\\.", ""), end = -2), Sats)]

comments[ , Sats := as.numeric(Sats)] 
```
#----------------------------------------------------------------------------------------------------------------------

# PART 2. Summary Table
### Divide in different periods
```{r}
### Divide in different periods
p_first_period = posts[Timestamp %between% c("2021-06-01","2021-12-31")]
p_second_period = posts[Timestamp %between% c("2022-01-01","2022-05-31")]
p_third_period = posts[Timestamp %between% c("2022-06-01","2022-12-31")]
p_fourth_period = posts[Timestamp %between% c("2023-01-01","2023-05-31")]
p_fifth_period = posts[Timestamp %between% c("2023-06-01","2023-12-31")]

c_first_period = comments[Timestamp %between% c("2021-06-01","2021-12-31")]
c_second_period = comments[Timestamp %between% c("2022-01-01","2022-05-31")]
c_third_period = comments[Timestamp %between% c("2022-06-01","2022-12-31")]
c_fourth_period = comments[Timestamp %between% c("2023-01-01","2023-05-31")]
c_fifth_period = comments[Timestamp %between% c("2023-06-01","2023-12-31")]
```

## Create the variables
### Periods
```{r}
Period = c("First", "Second", "Third", "Fourth", "Fifth")
```

### Unique Poster 
```{r}
Posters = c(
    uniqueN(p_first_period$Author),
    uniqueN(p_second_period$Author),
    uniqueN(p_third_period$Author),
    uniqueN(p_fourth_period$Author),
    uniqueN(p_fifth_period$Author))
```

### Total Sats from posts and from comments
```{r}
Post_Sats = c(
  sum(p_first_period$Sats,na.rm = TRUE),
  sum(p_second_period$Sats,na.rm = TRUE),
  sum(p_third_period$Sats,na.rm = TRUE),
  sum(p_fourth_period$Sats,na.rm = TRUE),
  sum(p_fifth_period$Sats,na.rm = TRUE))

Comment_Sats = c(
  sum(c_first_period$Sats,na.rm = TRUE),
  sum(c_second_period$Sats,na.rm = TRUE),
  sum(c_third_period$Sats,na.rm = TRUE),
  sum(c_fourth_period$Sats,na.rm = TRUE),
  sum(c_fifth_period$Sats,na.rm = TRUE))
```

### Only commentors
```{r}
f_commentors = anti_join(p_first_period, c_first_period, by = "Author")
s_commentors = anti_join(p_second_period, c_second_period, by = "Author")
t_commentors = anti_join(p_third_period, c_third_period, by = "Author")
f_commentors = anti_join(p_fourth_period, c_fourth_period, by = "Author")
fi_commentors = anti_join(p_fifth_period, c_fifth_period, by = "Author")
Only_commentors=c(nrow(f_commentors),nrow(s_commentors),nrow(t_commentors),nrow(f_commentors),nrow(fi_commentors))
```

### N posts and N comments
```{r}
N_posts=c(nrow(p_first_period),
  nrow(p_second_period),
  nrow(p_third_period),
  nrow(p_fourth_period),
  nrow(p_fifth_period))

N_comments=c(nrow(c_first_period),
  nrow(c_second_period),
  nrow(c_third_period),
  nrow(c_fourth_period),
  nrow(c_fifth_period))
```

### Type of post
```{r}
# link
N_link=c(nrow(p_first_period[Category == "link"]),
         nrow(p_second_period[Category == "link"]),
         nrow(p_third_period[Category == "link"]),
         nrow(p_fourth_period[Category == "link"]),
         nrow(p_fifth_period[Category == "link"]))

# discussion
N_discussion=c(nrow(p_first_period[Category == "discussion"]),
         nrow(p_second_period[Category == "discussion"]),
         nrow(p_third_period[Category == "discussion"]),
         nrow(p_fourth_period[Category == "discussion"]),
         nrow(p_fifth_period[Category == "discussion"]))

# bounty
N_bounty=c(nrow(p_first_period[Category == "bounty"]),
         nrow(p_second_period[Category == "bounty"]),
         nrow(p_third_period[Category == "bounty"]),
         nrow(p_fourth_period[Category == "bounty"]),
         nrow(p_fifth_period[Category == "bounty"]))
# poll

N_poll=c(nrow(p_first_period[Category == "poll"]),
         nrow(p_second_period[Category == "poll"]),
         nrow(p_third_period[Category == "poll"]),
         nrow(p_fourth_period[Category == "poll"]),
         nrow(p_fifth_period[Category == "poll"]))
         
```


## Creating the Summary Table
```{r}
summary_table=data.frame(Period,N_posts,N_comments,Posters,Post_Sats,Comment_Sats,Only_commentors,N_link,N_discussion,N_bounty,N_poll)
```

```{r}
summary_table
```
# LINK SUMMARY
### Check the links
```{r}
twitter=c(nrow(p_first_period[grepl("twitter", MainLink, ignore.case = TRUE)]),
          nrow(p_second_period[grepl("twitter", MainLink, ignore.case = TRUE)]),
          nrow(p_third_period[grepl("twitter", MainLink, ignore.case = TRUE)]),
          nrow(p_fourth_period[grepl("twitter", MainLink, ignore.case = TRUE)]),
          nrow(p_fifth_period[grepl("twitter", MainLink, ignore.case = TRUE)]))

youtube=c(nrow(p_first_period[grepl("youtube", MainLink, ignore.case = TRUE)]),
          nrow(p_second_period[grepl("youtube", MainLink, ignore.case = TRUE)]),
          nrow(p_third_period[grepl("youtube", MainLink, ignore.case = TRUE)]),
          nrow(p_fourth_period[grepl("youtube", MainLink, ignore.case = TRUE)]),
          nrow(p_fifth_period[grepl("youtube", MainLink, ignore.case = TRUE)]))

linkedin=c(nrow(p_first_period[grepl("linkedin", MainLink, ignore.case = TRUE)]),
          nrow(p_second_period[grepl("linkedin", MainLink, ignore.case = TRUE)]),
          nrow(p_third_period[grepl("linkedin", MainLink, ignore.case = TRUE)]),
          nrow(p_fourth_period[grepl("linkedin", MainLink, ignore.case = TRUE)]),
          nrow(p_fifth_period[grepl("linkedin", MainLink, ignore.case = TRUE)]))

bitcoinmagazine=c(nrow(p_first_period[grepl("bitcoinmagazine", MainLink, ignore.case = TRUE)]),
          nrow(p_second_period[grepl("bitcoinmagazine", MainLink, ignore.case = TRUE)]),
          nrow(p_third_period[grepl("bitcoinmagazine", MainLink, ignore.case = TRUE)]),
          nrow(p_fourth_period[grepl("bitcoinmagazine", MainLink, ignore.case = TRUE)]),
          nrow(p_fifth_period[grepl("bitcoinmagazine", MainLink, ignore.case = TRUE)]))

github=c(nrow(p_first_period[grepl("github", MainLink, ignore.case = TRUE)]),
          nrow(p_second_period[grepl("github", MainLink, ignore.case = TRUE)]),
          nrow(p_third_period[grepl("github", MainLink, ignore.case = TRUE)]),
          nrow(p_fourth_period[grepl("github", MainLink, ignore.case = TRUE)]),
          nrow(p_fifth_period[grepl("github", MainLink, ignore.case = TRUE)]))

yahoo=c(nrow(p_first_period[grepl("yahoo", MainLink, ignore.case = TRUE)]),
          nrow(p_second_period[grepl("yahoo", MainLink, ignore.case = TRUE)]),
          nrow(p_third_period[grepl("yahoo", MainLink, ignore.case = TRUE)]),
          nrow(p_fourth_period[grepl("yahoo", MainLink, ignore.case = TRUE)]),
          nrow(p_fifth_period[grepl("yahoo", MainLink, ignore.case = TRUE)]))
```

## Creating the Link Summary Table
```{r}
link_table=data.frame(Period,N_link, twitter,youtube,linkedin, bitcoinmagazine,github,yahoo)
```

```{r}
link_table
```
#------------------------------------------------------------------------------------------------------------------------
# PART 3. Network Analysis

## First Period
```{r}
first_graph <- copy(p_first_period)
first_graph <- first_graph[, .(Commentors = unlist(tstrsplit(CommentsItemCode, ", "))), by = "ItemCode"]
first_graph <- merge(first_graph, posts, by.x = 'ItemCode', by.y = 'ItemCode')
first_graph <- first_graph[, .(PostItemCode = ItemCode, CommentItemCode = Commentors, PostAuthor = Author)]
first_graph <- merge(first_graph, comments, by.x = 'CommentItemCode', by.y = 'ItemCode')
first_graph <- first_graph[, .(PostItemCode, CommentItemCode, PostAuthor, CommentAuthor = Author)]
first_graph <- first_graph[, .(PostAuthor, CommentAuthor)]
```

### Graph 

```{r}
g <- graph_from_data_frame(first_graph, directed = F)
```

```{r}
g1=igraph::simplify(g, remove.multiple = T, remove.loops = T)
```

```{r}
plot( g1,vertex.label=NA,
      edge.width = 1,
      edge.arrow.width = 0.3,
      edge.arrow.size = 0.5,
      vertex.label.cex = 1,
      asp = 0.35,
      margin = -0.1)
```


## Second Period
```{r}
second_graph <- copy(p_second_period)
second_graph <- second_graph[, .(Commentors = unlist(tstrsplit(CommentsItemCode, ", "))), by = "ItemCode"]
second_graph <- merge(second_graph, posts, by.x = 'ItemCode', by.y = 'ItemCode')
second_graph <- second_graph[, .(PostItemCode = ItemCode, CommentItemCode = Commentors, PostAuthor = Author)]
second_graph <- merge(second_graph, comments, by.x = 'CommentItemCode', by.y = 'ItemCode')
second_graph <- second_graph[, .(PostItemCode, CommentItemCode, PostAuthor, CommentAuthor = Author)]
second_graph <- second_graph[, .(PostAuthor, CommentAuthor)]
```

### Graph 
```{r}
g2 <- graph_from_data_frame(second_graph, directed = F)
```

```{r}
g2=igraph::simplify(g2, remove.multiple = T, remove.loops = T)
```

```{r}
plot( g2,vertex.label=NA,
      edge.width = 1,
      edge.arrow.width = 0.3,
      edge.arrow.size = 0.5,
      vertex.label.cex = 1,
      asp = 0.35,
      margin = -0.1)
```

## Third Period
```{r}
third_graph <- copy(p_third_period)
third_graph <- third_graph[, .(Commentors = unlist(tstrsplit(CommentsItemCode, ", "))), by = "ItemCode"]
third_graph <- merge(third_graph, posts, by.x = 'ItemCode', by.y = 'ItemCode')
third_graph <- third_graph[, .(PostItemCode = ItemCode, CommentItemCode = Commentors, PostAuthor = Author)]
third_graph <- merge(third_graph, comments, by.x = 'CommentItemCode', by.y = 'ItemCode')
third_graph <- third_graph[, .(PostItemCode, CommentItemCode, PostAuthor, CommentAuthor = Author)]
third_graph <- third_graph[, .(PostAuthor, CommentAuthor)]
```

### Graph 
```{r}
g3 <- graph_from_data_frame(third_graph, directed = F)
```

```{r}
g3=igraph::simplify(g3, remove.multiple = T, remove.loops = T)
```

```{r}
plot( g3,vertex.label=NA,
      edge.width = 1,
      edge.arrow.width = 0.3,
      edge.arrow.size = 0.5,
      vertex.label.cex = 1,
      asp = 0.35,
      margin = -0.1)
```

### Fourth Period
```{r}
fourth_graph <- copy(p_fourth_period)
fourth_graph <- fourth_graph[, .(Commentors = unlist(tstrsplit(CommentsItemCode, ", "))), by = "ItemCode"]
fourth_graph <- merge(fourth_graph, posts, by.x = 'ItemCode', by.y = 'ItemCode')
fourth_graph <- fourth_graph[, .(PostItemCode = ItemCode, CommentItemCode = Commentors, PostAuthor = Author)]
fourth_graph <- merge(fourth_graph, comments, by.x = 'CommentItemCode', by.y = 'ItemCode')
fourth_graph <- fourth_graph[, .(PostItemCode, CommentItemCode, PostAuthor, CommentAuthor = Author)]
fourth_graph <- fourth_graph[, .(PostAuthor, CommentAuthor)]
```

### Graph 
```{r}
g4 <- graph_from_data_frame(fourth_graph, directed = F)
```

```{r}
g4=igraph::simplify(g4, remove.multiple = T, remove.loops = T)
```

```{r}
plot( g4,vertex.label=NA,
      edge.width = 1,
      edge.arrow.width = 0.3,
      edge.arrow.size = 0.5,
      vertex.label.cex = 1,
      asp = 0.35,
      margin = -0.1)
```

### Fifth Period
```{r}
fifth_graph <- copy(p_fifth_period)
fifth_graph <- fifth_graph[, .(Commentors = unlist(tstrsplit(CommentsItemCode, ", "))), by = "ItemCode"]
fifth_graph <- merge(fifth_graph, posts, by.x = 'ItemCode', by.y = 'ItemCode')
fifth_graph <- fifth_graph[, .(PostItemCode = ItemCode, CommentItemCode = Commentors, PostAuthor = Author)]
fifth_graph <- merge(fifth_graph, comments, by.x = 'CommentItemCode', by.y = 'ItemCode')
fifth_graph <- fifth_graph[, .(PostItemCode, CommentItemCode, PostAuthor, CommentAuthor = Author)]
fifth_graph <- fifth_graph[, .(PostAuthor, CommentAuthor)]
```

### Graph 
```{r}
g5 <- graph_from_data_frame(fifth_graph, directed = F)
```

```{r}
g5=igraph::simplify(g5, remove.multiple = T, remove.loops = T)
```

```{r}
plot( g5,vertex.label=NA,
      edge.width = 1,
      edge.arrow.width = 0.3,
      edge.arrow.size = 0.5,
      vertex.label.cex = 1,
      asp = 0.35,
      margin = -0.1)
```